Structured Versus Decorated Cospans

### 2021-06-22 cospans_new.tex
edits

We now compare structured and decorated cospans.   In \cref{thm:decorated_cospans_1} we built a double category of decorated cospans from a bifibration $\Phi \maps \A/U \to \A$.    In \cref{thm:structured_cospans}, we built a double category of structured cospans from a bifibration $\Psi \maps L/\X \to \A$ in a very similar way.   We now show that under certain conditions $L$ is left adjoint to $U$.   Whenever this happens, $\A/U$ is isomorphic to $L/\X$, by a simple and general fact about arrow categories.    The bifibrations $\Phi$ and $\Psi$ are also isomorphic.   By the functoriality of Shulman's construction in \cref{lem:shulman}, it follows that the double category of decorated cospans is isomorphic to the double category of structured cospans.

The key issue is thus to determine when $L$ is left adjoint to $U$.  For this, let $\Rex$ be the 2-category of categories with finite colimits, functors preserving finite colimits, and natural transformations.  Also let $\SMC$ be the 2-category of symmetric monoidal categories, strong symmetric monoidal functors and natural transformations.   Recall that for us a category $\C \in \Rex$ comes with a choice of finite colimits, so it gives a specific cocartesian monoidal category $(\C,+)$, and this induces a 2-functor $\Rex \to \SMC$.

@@

To show that the symmetric monoidal double categories $F\lCsp$ and $_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans} are isomorphic, we make use of a result of Shulman \cite[Theorem~14.11]{Shulman2010}.   Namely, he showed that the $\mathbb{F}\mathbf{r}$ construction of \cref{lem:shulman} extends to a 2-functor 
\[ \mathbb{F}\mathbf{r} \maps \mathbf{FMF} \to \mathbf{FDbl} \]
where $\mathbf{FMF}$ is the 2-category consisting of
\begin{itemize}
\item Beck--Chevalley monoidal bifibrations,
\item strong monoidal morphisms of bifibrations, and
\item monoidal transformations of bifibrations
\end{itemize}
while $\mathbf{FDbl}$ is the 2-category of 
\begin{itemize}
\item fibrant double categories, 
\item double functors, and 
\item double transformations. 
\end{itemize}
Given two monoidal bifibrations $\Phi \maps \A \to \mathsf{B}$ and $\Phi' \maps \A' \to \mathsf{B}'$, a \define{strong monoidal morphism of bifibrations} from $\Phi$ to $\Phi'$ consists of a pair of strong monoidal functors $F_1$ and $F_2$ making the following square commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\A'$};
\node (C) at (0,-1) {$\mathsf{B}$};
\node (D) at (1,-1) {$\mathsf{B}'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F_1$} (B)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[above]{$F_2$} (D);
\end{tikzpicture}
\]
and such that the identity $\Phi' F_1 = F_2 \Phi$ is a monoidal transformation. The two monoidal bifibrations in question are the functors $\Phi$ and $\Psi$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively, which share $\A$ as a common base, and thus we only need a single strong monoidal functor $F \maps \A/U \to L/\X$ making this diagram commute:
\[
\begin{tikzpicture}[scale=1.0]
\node (A) at (0,0) {$\A/U$};
\node (B) at (2,0) {$L/ \X$};
\node (C) at (1,-1.5) {$\A$};
%\node (D) at (2,-1) {$\A$};
%\node (E) at (1,0) {$\Downarrow \alpha_1$};
%\node (F) at (1,-1) {$\Downarrow \alpha_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
%(B) edge [bend left] node[above]{$$} (A)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Psi$} (C);
%(C) edge [bend left] node[above]{$G_1$} (C)
%(C) edge [bend right] node[above]{$G_2$} (C);
\end{tikzpicture}
\]
{\john with $\Psi F=\Phi$ a monoidal transformation.}  There is an isomorphism $F \maps \A/U \to L/\X$ arising from the fact that $U$ is right adjoint to $L$.   It is clear that $F$ is strong monoidal via componentwise binary coproducts which are preserved by both $L$ and $U$, {\john and that the identity $\Psi F=\Phi$ is a monoidal transformation.} As $F$ is an isomorphism, we obtain an isomorphism 
\[  \mathbb{F}\mathbf{r}(F) \maps \mathbb{F}\mathbf{r}(\Phi) \to \mathbb{F}\mathbf{r}(\Psi)\] between the two monoidal double categories $\mathbb{F}\mathbf{r}(\Phi)=F\lCsp$ and $\mathbb{F}\mathbf{r}(\Psi)= {}_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.  {\john One can check by hand that this isomorphism 
is symmetric monoidal.}
\end{proof}

>>>>

We now compare structured and decorated cospans.   In \cref{thm:decorated_cospans_1} we built a double category of decorated cospans from a bifibration $\Phi \maps \A/U \to \A$.    In \cref{thm:structured_cospans}, we built a double category of structured cospans from a bifibration $\Psi \maps L/\X \to \A$ in a very similar way.   We now show that under certain conditions $L$ is left adjoint to $U$.   Whenever this happens, $\A/U$ is isomorphic to $L/\X$, by a simple and general fact about arrow categories.    The bifibrations $\Phi$ and $\Psi$ are also isomorphic.   By the functoriality of Shulman's construction in \cref{lem:shulman}, {\chris but this functoriality is not spelled out there at the moment! We should probably add it, see next next comment.} it follows that the double category of decorated cospans is isomorphic to the double category of structured cospans.

The key issue is thus to determine when $L$ is left adjoint to $U$.  For this, let $\Rex$ be the 2-category of categories with finite colimits, functors preserving finite colimits, and natural transformations.  Also let $\SMC$ be the 2-category of symmetric monoidal categories, strong symmetric monoidal functors {\chris symmetric strong mon funs?} and natural transformations.   Recall that for us a category $\C \in \Rex$ comes with a choice of finite colimits, so it gives a specific cocartesian monoidal category $(\C,+)$, and this induces a 2-functor $\Rex \to \SMC$.

@@

To show that the symmetric monoidal double categories $F\lCsp$ and $_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans} are isomorphic, we make use of a result of Shulman \cite[Theorem~14.11]{Shulman2010}.   {\chris Shouldn't this material be moved outside the proof environment, probably earlier?} Namely, he showed that the $\mathbb{F}\mathbf{r}$ construction of \cref{lem:shulman} extends to a 2-functor 
\[ \mathbb{F}\mathbf{r} \maps \mathbf{FMF} \to \mathbf{FDbl} \]
where $\mathbf{FMF}$ is the 2-category consisting of
\begin{itemize}
\item Beck--Chevalley monoidal bifibrations,
\item strong monoidal morphisms of bifibrations, and
\item monoidal transformations of bifibrations
\end{itemize}
while $\mathbf{FDbl}$ is the 2-category of 
\begin{itemize}
\item fibrant double categories, 
\item double functors, and 
\item double transformations. 
\end{itemize}
Given two monoidal bifibrations $\Phi \maps \A \to \mathsf{B}$ and $\Phi' \maps \A' \to \mathsf{B}'$, a \define{strong monoidal morphism of bifibrations} from $\Phi$ to $\Phi'$ consists of a pair of strong monoidal functors $F_1$ and $F_2$ making the following square commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\A'$};
\node (C) at (0,-1) {$\mathsf{B}$};
\node (D) at (1,-1) {$\mathsf{B}'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F_1$} (B)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[above]{$F_2$} (D);
\end{tikzpicture}
\]
and such that the identity $\Phi' F_1 = F_2 \Phi$ is a monoidal transformation. The two monoidal bifibrations in question are the functors $\Phi$ and $\Psi$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively, which share $\A$ as a common base, and thus we only need a single strong monoidal functor $F \maps \A/U \to L/\X$ making this diagram commute:
\[
\begin{tikzpicture}[scale=1.0]
\node (A) at (0,0) {$\A/U$};
\node (B) at (2,0) {$L/ \X$};
\node (C) at (1,-1.5) {$\A$};
%\node (D) at (2,-1) {$\A$};
%\node (E) at (1,0) {$\Downarrow \alpha_1$};
%\node (F) at (1,-1) {$\Downarrow \alpha_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
%(B) edge [bend left] node[above]{$$} (A)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Psi$} (C);
%(C) edge [bend left] node[above]{$G_1$} (C)
%(C) edge [bend right] node[above]{$G_2$} (C);
\end{tikzpicture}
\]
{\john with $\Psi F=\Phi$ a monoidal transformation.}  There is an isomorphism $F \maps \A/U \to L/\X$ arising from the fact that $U$ is right adjoint to $L$.   It is clear that $F$ is strong monoidal via componentwise binary coproducts which are preserved by both $L$ and $U$, {\john and that the identity $\Psi F=\Phi$ is a monoidal transformation.} As $F$ is an isomorphism, we obtain an isomorphism 
\[  \mathbb{F}\mathbf{r}(F) \maps \mathbb{F}\mathbf{r}(\Phi) \to \mathbb{F}\mathbf{r}(\Psi)\] between the two monoidal double categories $\mathbb{F}\mathbf{r}(\Phi)=F\lCsp$ and $\mathbb{F}\mathbf{r}(\Psi)= {}_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.  {\john One can check by hand that this isomorphism 
is symmetric monoidal.}
{\chris Even so, shouldn't we also add some of the explicit isomorphism mapping to make the isomorphism between those two double categories more concrete?}
\end{proof}
